Understanding the Socio-Economic Determinants of Voter Behavior: An Analysis of the 2020 U.S. Presidential Election

### 2024-03-12 paper/us_election_2020.qmd
complete intro

# Introduction
Four years ago, the United States once again was at the forefront of an important electoral contest between the Democratic and Republican parties in the 2020 presidential election. The Democratic and Republican parties are represented by Donald Trump and Joe Biden respectively. Donald Trump achieved a remarkable victory in the 2016 election against Hillary Clinton. His winning was surprising and it led people to challenge the preconceived notions and reliability of pre-election polling data [citation needed]. On the other hand, Joe Biden, the former vice-president under Barack Obama's administration from 2008 to 2016, led the Democratic campaign [citation needed]. He aimed to regain control from the Republican party. In the end, Biden won the election by 51.3% majority votes [citation needed]. It is important for us to understand how various demographic, economic, and social factors may influence individuals' voting decisions, which is why we decide to analyze the 2020 election data set and build a prediction model. This allows us to discover trends and address potential issues in society to better reflect the needs and preferences of diverse communities.

In this paper, a logistic regression model is used to forecast the outcome of the 2020 election, with data from the Cooperative Election Study (CES)  [citation needed]. Logistic regression is a great choice since it is used to predict binary outcomes, such as election results (Trump or Biden). Our analysis focuses on estimating the likelihood of victory for either Trump or Biden, based on various demographic, geographic, and socioeconomic factors captured in the CES data set, which are race, region that they live, and employment status.

**add： 4) what was done; 5) what was found after we finish the results section

The remainder of this paper is structured into different sections. @sec-data demonstrates the data used for our report and includes some tables and graphs to illustrate the different groups of people in our data. @sec-model builds the model and discusses its justification and explanation. @sec-result highlights the results of the predictions using tables and graphs. @sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, and employment status.

## Estimand
real vote_for biden or trump

>>>>

# Introduction
Four years ago, the United States once again was at the forefront of an important electoral contest between the Democratic and Republican parties in the 2020 presidential election. The Democratic and Republican parties are represented by Donald Trump and Joe Biden respectively. Donald Trump achieved a remarkable victory in the 2016 election against Hillary Clinton. On the other hand, Joe Biden, the former vice-president under Barack Obama's administration from 2008 to 2016, led the Democratic campaign. He aimed to regain control from the Republican party. In the end, Biden won the election by 51.3% majority votes (@cnn_2020), which Trump questioned its credibility. Due to the COVID-19 pandemic in 2020, the vote system changed to in-mail voting instead of in-person voting, which was why Trump believed that some votes were stolen (@trump_stolenvote). Therefore, it is important for us to analyze why people voted for Biden and Trump in 2020 election and how various demographic, economic, and social factors may influence individuals' voting decisions. Perhaps then we could see why Biden eventually won. We analyze the 2020 election data set from Cooperative Election Study (CES) [citation needed] and build a prediction model for individual voting preference. This allows us to discover trends and address potential political, economic, and societal issues to better reflect the needs and preferences of diverse communities.

In this paper, a logistic regression model is used to forecast the outcome of the 2020 election, with data from the Cooperative Election Study. Logistic regression is a great choice since it is used to predict binary outcomes, such as election results (Trump or Biden). Our analysis focuses on estimating the likelihood of victory for either Trump or Biden, based on various demographic, geographic, and socioeconomic factors captured in the CES data set, which are race, region that they live, and employment status. The estimand in this paper is the number of people who support Trump and Biden in reality. However, it is difficult to measure the exact number of people who support Trump and Biden since there are millions of people in the United States and not all of them with the rights to vote would actually vote. Therefore, in this paper, we attempt to estimate the estimand using a logistic regression model which is trained using 2020 election data set from CES.

We found that in the 2020 presidential election, Biden had strong support from racial groups including black, Hispanic, and Asian populations. Geographically, Biden gained greater support in the Northeast and West regions, while Trump secured more votes in the South. Additionally, both full-time and part-time workers, as well as students and those facing unemployment or layoffs, leaned towards Biden.

The remainder of this paper is structured into different sections. @sec-data demonstrates the data used for our report and includes some tables and graphs to illustrate the different groups of people in our data. @sec-model builds the model and discusses its justification and explanation. @sec-result highlights the results of the predictions using tables and graphs. @sec-discussion contains discussions that conducted based on the findings, which addresses the voting prediction results based on race, region, employment status, and the influence of COVID-19 and in-mail voting systems.
